### 2013-09-25 manuscripts/dpp-msbayes/dpp-msbayes.tex
More work on model description for dpp ms.

@@

\begin{linenomath}
The total sample space of ordered realizations of \divTimeMapVector is all the
possible partitions of \npairs{} elements into 1 to \npairs{} categories.
The total number of possible partitions is a sum of the Stirling numbers of
the second kind over all possible number of categories \divTimeNum
\begin{equation}
    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
    \label{eq:bell}
\end{equation}
which is the Bell Number \citep{Bell1934}.
The original \msb model only samples over the unordered realizations of
\divTimeIndexVector, such that the sample space is reduced to all the possible
integer partitions of \npairs{} \citep{Oaks2012,OeisPartitionNumber,
    OeisPartitionTriangle,Malenfant2011}.
This has the advantage of reducing the sample space of the number of
divergence models, but has two disadvantages.
First, additional information about the data is lost.
By throwing away the identity of the \npairs{} pairs of populations, all
information regarding the amount of data (e.g., the number gene copies
collected from each of the populations [\popSampleSize{1}{} and
\popSampleSize{2}{}], the number of loci, and the length of the loci), the
information content in the data (i.e., the amount of sequence variation), as
well as the taxon- and locus-specific parameters (\hkyModel{}{},
\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
\locusMutationRateScalar{}) for each pair is lost.
The original descriptions of the \msb model claim that the independence
between the sample sizes \popSampleSize{1}{}, \popSampleSize{2}{} and
the summary statistic $\pi_b$ calculated from the alignment for each
population pair allows exchangeability.
Second, the identity of the pairs of populations is lost.
\end{linenomath}

>>>>

\begin{linenomath}
To avoid calculating the likelihood terms of Equation \ref{eq:fullModelCompact},
we distill each sequence alignment \alignment{}{} into a vector of insufficient
summary statistics \alignmentSS{}{}, thus replacing the full dataset
$\alignmentVector = (\alignment{1}{1}, \ldots,
    \alignment{\npairs{}}{\nloci{\npairs{}}})$
with vectors of summary statistics for each alignment
$\ssVector{} = (\alignmentSS{1}{1}, \ldots,
    \alignmentSS{\npairs{}}{\nloci{\npairs{}}})$.
This allows us to estimate the approximate joint posterior distribution
{\small
\begin{equation}
    p(\geneTreeVector, \divTimeMapVector, \demographicParamVector, 
    \locusMutationRateScalarVector, \locusRateHetShapeParameter \given
    \ssSpace, \hkyModelVector, \ploidyScalarVector,
    \mutationRateScalarConstantVector) =
    \frac{p(\ssSpace \given \geneTreeVector, \hkyModelVector)
        p(\geneTreeVector \given \divTimeMapVector, \demographicParamVector,
        \locusMutationRateScalarVector, \ploidyScalarVector,
        \mutationRateScalarConstantVector)
        p(\locusMutationRateScalarVector \given \locusRateHetShapeParameter)
        p(\locusRateHetShapeParameter)
        p(\divTimeMapVector)
        p(\demographicParamVector)}{
        p(\ssSpace)},
    \label{eq:approxModelCompact}
\end{equation}
}
where \ssSpace is the multidimensional Euclidean space around the vector of
summary statistics, the radius of which is the tolerance \tol.
The approximation of this posterior is caused by the insufficiency of the
statistics and the \tol being greater than zero.
We describe the full model in detail before delving into the numerical
method of estimating the approximate model.
\end{linenomath}

@@

\begin{linenomath}
The total sample space of ordered realizations of \divTimeMapVector is all the
possible partitions of \npairs{} elements into 1 to \npairs{} categories.
The total number of possible partitions is a sum of the Stirling numbers of
the second kind over all possible number of categories \divTimeNum
\begin{equation}
    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
    \label{eq:bell}
\end{equation}
which is the Bell Number \citep{Bell1934}.
The original \msb model only samples over the unordered realizations of
\divTimeIndexVector, such that the sample space is reduced to all the possible
integer partitions of \npairs{} \citep{Oaks2012,OeisPartitionNumber,
    OeisPartitionTriangle,Malenfant2011}.
This has the advantage of reducing the sample space of the number of
divergence models, but has two disadvantages.
First, additional information about the data is lost.
By throwing away the identity of the \npairs{} pairs of populations, all
information regarding the amount of data (e.g., the number gene copies
collected from each of the populations [\popSampleSize{1}{} and
\popSampleSize{2}{}], the number of loci, and the length of the loci), the
information content in the data (i.e., the amount of sequence variation), as
well as the taxon- and locus-specific parameters (\hkyModel{}{},
\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
\locusMutationRateScalar{}) for each pair is lost.
The original descriptions of the \msb model claim that the independence
between the sample sizes \popSampleSize{1}{}, \popSampleSize{2}{} and
the summary statistic $\pi_b$ calculated from the alignment for each
population pair allows exchangeability.
This is incorrect.
The other summary statistics, and the alignments they are based on are
not independent of the sample sizes.
The alignments and summary statitics are not identically distributed because of
taxon- and locus-specific parameters \hkyModel{}{},
\mutationRateScalarConstant{}{}, and \ploidyScalar{}{}
Second, the identity of the pairs of populations is lost.
\end{linenomath}
